Reduced GW invariants from cuspidal curves

### 2018-01-09 draft_merged.tex
notazione in Sec.3 e localised virtual pullback commuta con pushforward

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem*{teo*}{Theorem}
\newtheorem{ipotesi}{ipotesi}
\newtheorem*{nota}{Nota}
\newtheorem{claim}{Claim}
\newtheorem{question}[thm]{Question}

@@

\begin{dfn}
Let $\mathfrak M_{1,n}^{\operatorname{wt}=d,\text{st}}(1)$ be the stack of \emph{at worst cuspidal} projective reduced connected $n$-marked curves of arithmetic genus $1$ that are weighted-stable with total weight $d$, i.e. the weight is nonnegative, and every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component  of weight $0$ has at least \emph{two} special points).
\end{dfn}

We shall denote by $\hat{\pi}\colon \widehat{\cC}\to \MM_{1,n}(1)$ the universal curve (and use similar notation for all its pullbacks). We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$ of $\widehat{\cC}\to \mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}(1)$ with universal line bundle $\widehat{\mathcal{L}}$. We shall often use shorthand notation $\widehat{\MM}$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}(1)$ and $\widehat{\mathfrak P}$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$. Let $\hat{\PtoM}$ denote the projection $\widehat{\mathfrak P}\to \widehat{\MM}$.

$\dvr$ shall always denote the spectrum of a discrete valuation ring (DVR) with closed point $0$ and generic point $\eta$.

@@

\section{From $p$-fields to the quintic threefold}
We introduce the moduli space $\Mone{1}{\PP^4}{d}^p$ of $1$-stable curves with $p$-fields, we endow it with a cosection-localised $0$-dimensional virtual class supported on a proper substack (depending on the polynomial $\w$), and we show that its degree coincides with the genus $1$ cuspidal invariants of the quintic threefold $X$ up to a sign. This is a word-by-word repetition of the arguments in \cite{CLpfields} once noticed that they carry over to the situation where we work with a family of Gorenstein (not necessarily nodal) curves. Our aim in this section is to improve the legibility of this paper by providing the non-expert reader with a r\'{e}sum\'{e} of some of the key ideas contained in \cite{CLpfields}; it can otherwise be skipped.
\subsection{Moduli of sections}
\begin{dfn}
 Let $B$ be an algebraic stack and let $\pi\colon \cC\to B$ be a flat proper morphism of finite presentation which is representable by algebraic spaces and whose geometric fibers are reduced l.c.i. curves. Let $\mathcal Z$ be an algebraic stack smooth representable and quasi-projective over $\cC$. The \emph{cone of sections} of $\mathcal Z$ over $\cC$ is a $B$-stack $\mathfrak S$ defined by:
 \[\mathfrak{S}(S\to B)=\left\{\text{sections of}\ \mathcal Z_S\to\cC_S\right\};\]

@@

\begin{lem}\label{lemma:def_X}
 There is a locally closed substack $\mathcal X\subseteq\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ representing morphisms $C\to \widehat{C}$ that contract elliptic tails of weight $0$ to cusps and are weight-preserving isomorphisms everywhere else. 
\end{lem}

\begin{lem}\label{lamma:projection_iso}
 The first projection $\pr_1\colon \operatorname{Mor}_{\MM\times\hM}(\cC,\hC)\to\MM$ restricted to $\mathcal X$ is an isomorphism with $\MM$.
\end{lem}

\begin{proof}\ref{lemma:def_X}
Recall that $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ is an algebraic stack; in fact the map to $\MM\times\hM$ is representable by algebraic spaces \cite{OlssonHOM}. Let $\phi\colon\cC\to\hC$ denote the universal morphism. We now proceed to construct $\mathcal X$ as a locally closed substack in the space of morphisms.

@@

\item[Step 1] We first lift the problem to the level of Picard stacks in order to transmute the property of being weight-preserving into the more manageable one of preserving the line bundles. Recall the notation:
 \[\PtoM\colon\mathfrak P=\mathfrak{Pic}^{tot deg=d,st}_{1,n}\to\MM,\quad \hat\PtoM\colon\hP=\mathfrak{Pic}^{tot deg=d,st}_{1,n}(1)\to\hM\]
  for the Picard stacks of $\pi$ and $\hat \pi$ respectively, with universal line bundles $\L$ and $\hL$. The projections $\PtoM$ and $\hat \PtoM$ are defined by taking the multi-degree of a line bundle. We can now look at the algebraic stack $\operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$ with universal morphism
 $\Phi$ and natural projection $\Pi$ to $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$. There exists a locally closed substack $\mathcal{Y'}\subseteq \operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$ representing those morphisms that \emph{preserve the line bundles}. Indeed, given a smooth chart $S\to\operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$, the locus of $s\in S$ where $\Phi_s^*\hL_s\cong\mathcal{L}_s$ is nothing but the locus $T$ where the two sections $\mathcal{L}_S$ and $\Phi_S^*\hL_S$ of $\mathfrak{P}(S)\to\MM(S)$ are isomorphic. In other words we are looking at the fiber product 
  \[\xymatrix{T\ar[r]\ar[d]&\mathfrak{P}\ar[d]^{\Delta}\\
  S\ar[r]&\mathfrak{P}\times_{\MM}\mathfrak{P}}\]
Being $\mathfrak{P}\to\MM$ representable by locally separated algebraic spaces \cite[Theorem~8.3.1]{neron}, $\Delta$ is a quasi-compact locally closed immersion \cite[\href{http://stacks.math.columbia.edu/tag/04YU}{Tag 04YU}]{stacks-project}, so in particular $T\subseteq S$ is locally closed.

\item[Step 2] 
Furthermore there is a closed substack $\mathcal Y\subseteq \mathcal Y'$ representing \emph{surjective morphisms that preserve the markings}.

@@

\item[Step 3] We may now get back to $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$. Let $\mathcal{X}'$ be the image of $\mathcal Y$ under $\Pi$. This is a constructible substack of $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ by Chevalley's theorem \cite[Theorem 5.9.4]{LMB}. Recall that to show that a constructible set is open (respectively closed) it is enough to check that it contains all the generisations of its points (respectively all the specialisations) \cite[\href{http://stacks.math.columbia.edu/tag/0DQN}{Tag 0DQN}\href{http://stacks.math.columbia.edu/tag/0903}{Tag 0903}]{stacks-project}. Finally, under Noetherian assumptions, two points related by specialisation/generisation are contained in the image of a DVR scheme \cite[\href{http://stacks.math.columbia.edu/tag/054F}{Tag 054F}]{stacks-project}.

@@

\begin{enumerate}
\item\emph{$\phi$ contracts only components of weight $0$.} We show that this is open. Assume that $\dvr$ is a DVR scheme with closed point $0$ and generic point $\eta$, and we are given $\dvr\to\mathcal X'$ such that $\phi_0\colon\cC_0\to \hC_0$ does not contract any component of positive weight. Suppose there exists an irreducible component $D_\eta\subseteq \cC_{\eta}$ of positive weight $d_D$ which is contracted by $\phi_{\eta}.$ The contracted locus, i.e. $\left\{c\in \cC_{\dvr} | \dim_c\phi^{-1}(\phi(c))\geq 1\right\}$, is closed by semicontinuity of fiber dimension, hence it contains all the components $D_i\subseteq \cC_0$ to which $D_\eta$ specialises. At least one of them has positive weight, since the sum of their weights is $d_D$, which is a contradiction.

\item\emph{$\phi$ has degree 1 on every non contracted component} or, equivalently, for every $S\to\mathcal X'$ there is an $S$-dense open in $\hC_S$ such that the restriction of $\phi_S$ to its preimage is an isomorphism.
This is an open and closed condition; we show it is open. Let $\dvr$ be a DVR scheme as above and assume that $\phi_0$ satisfies the property. Since $\phi_\dvr$ is proper, we may consider 
\[\phi_{\dvr,*}[\cC_{\dvr}]=\sum n_i[\hC_{\dvr,i}]\in A_2(\hC_\dvr).\] Applying Gysin pull-back to $0$ (which is a regular closed point of the base) \cite[Prop. 10.1(a)]{FUL}, we see that all the $n_i$'s are 1 for those $\hC_i$'s such that $0^![\hC_i]\neq 0$. On the other hand there is no irreducible component of $\hC$ supported on $\hC_\eta$.

@@

\item[Step 4] We have imposed that, if $\phi$ contracts a subcurve $E$ of the fiber, it must have weight $0$. Since the fibers of $\hC$ only have nodes and cusps as singularities, and the markings are required to be smooth points, we observe that $E$ must be unmarked and have arithmetic genus $1$ by weighted stability; furthermore $\left|E\cap\overline{C\setminus E}\right|\leq 2$, i.e. $E$ is either an elliptic tail or an elliptic bridge. There are basically two possibilities:
\begin{enumerate}
\item $\phi$ contracts an elliptic tail to a cusp and is an isomorphism everywhere else, or there is no elliptic tail to start with and $\phi$ is an isomorphism;
\item the elliptic tail/bridge is contracted to a smooth point/node, then a non-separating node or a cusp must appear somewhere else in the fiber of $\hC$ in order to preserve the arithmetic genus.
\end{enumerate} 
We want to avoid the second scenario, so we define the open substack $\mathcal X\subseteq \mathcal X'$ as follows.
Given $\mathcal{C}_S\to\hC_S\in\mathcal X'(S)$, let $U\subseteq \hC$ be the \emph{maximal} $S$-dense open subset such that $\phi_S|_{\phi_S^{-1}(U)}\colon\phi_S^{-1}(U)\to U$ is an isomorphism, and $V$ its closed complement in $\hC_S$. 
Then $\mathcal X$ is the open locus \cite[\href{http://stacks.math.columbia.edu/tag/055G}{Tag 055G}]{stacks-project} where the fibers of $\pi|_{\phi^{-1}(V)}\colon \phi^{-1}(V)\to S$ are geometrically connected.
\end{description} 
This concludes the construction of the locally closed substack $\mathcal X\subseteq \operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$.
 \end{proof}

 \begin{proof}\ref{lamma:projection_iso}
 This result will follow from an application of Zariski's Main Theorem for algebraic spaces.
First we claim that the projection $\pr_{1|\mathcal X}\colon\mathcal X\to \MM$ is \emph{representable by algebraic spaces}: by \cite[\href{http://stacks.math.columbia.edu/tag/04Y5}{Tag 04Y5}]{stacks-project} we only need to check that it is faithful, and by \cite[Theorem 2.2.5]{CONR} it is enough to look at geometric points. Hence we need to say that, given $\phi\colon C\to \widehat{C}$ a $K$-point of $\X$, we have
$\operatorname{Aut}(\phi)\subseteq\operatorname{Aut}(C)$. Recall that automorphisms of $\phi$ are commutative diagrams:
\bcd C\ar[r,"\phi"]\ar[d,"\psi"] & \widehat{C}\ar[d,"\hat\psi"] \\ C\ar[r,"\phi"] & \widehat{C} \ecd
 Now $\hat\psi$ is determined by $\psi$ due to how $\phi$ behaves (contracting elliptic tails and being an isomorphism everywhere else).

Secondly $\pr_{1|\mathcal X}$ is \emph{proper}: this can be seen using the valuative criterion:
\bcd
\eta'=\operatorname{Spec}(K')\ar[r]\ar[d] & \eta=\operatorname{Spec}(K)\ar[r]\ar[d] & \mathcal X \ar[d] \\
\dvr'=\operatorname{Spec}(R')\ar[r]\ar[urr,dashed,"\exists?", left] & \dvr=\operatorname{Spec}(R)\ar[r] & \MM
\ecd
Let $\pi \colon \cC_\dvr\to \dvr$ be the family of nodal curves on $\dvr$; there are three cases to consider:
\begin{enumerate}[label=(\alph*)]
\item the central fiber contains no elliptic tail, then the same is true for $\cC_\eta$, hence $\phi_\eta$ is an isomorphism. We can extend $\phi_\eta$ as follows:
\bcd
\cC_\eta\arrow{r}{\sim}[swap]{\phi_\eta}\ar[d,hook,"\iota"] & \hC_\eta\ar[d,hook,"\iota\circ\phi_\eta^{-1}"] \\
\cC_\dvr\ar[r,"\id_\cC"] & \cC_\dvr=:\hC_\dvr
\ecd
Another extension $\phi'\colon \cC_\dvr\cong \hC_\dvr$ would be isomorphic to the previous one via:
\bcd
\cC_\dvr\ar[r,"\phi'"]\ar[d,"\id"] &\hC_\dvr\ar[d,"(\phi')^{-1}"] \\
\cC_\dvr\ar[r,"\id"] & \cC_\dvr
\ecd

\end{enumerate}
If instead $\cC_0$ has got an elliptic tail, then we have two possibilities:
\begin{enumerate}[label=(\alph*)]
\setcounter{enumi}{1}
\item $\cC_{\eta}$ has got an elliptic tail as well; that is the image of $\dvr\to\MM$ is contained in the boundary, so we can find a lift
\bcd
 & \MM_{1,1}\times\MM_{0,1+n}^{\rm{wt}=d,\rm{st}}\ar[d] & \\
 \dvr\ar[r]\ar[ur,dashed] & \mathfrak D_{\{1,\emptyset\},\{0,n\}}\ar[r,hook] &\MM.
\ecd
 Then $\cC_\dvr$ is the pushout of a family of rational curves $\mathcal R_\dvr$ and a family of genus one curves $\mathcal Z_\dvr$:
%The map $S\to\partial \mathfrak M_{1,n}$ lifts to the normalisation of the boundary, i.e. $\mathfrak M_{0,n_0+1}\times\mathfrak M_{1,n_1+1}$, which gives two families of pointed curves on $S$; there is a map from the pushout to $\mathcal C_S$ by the universal property, and it is proper bijective and flat (crit\`{e}re de platitude par fibres \cite[\href{http://stacks.math.columbia.edu/tag/039A}{Tag 039A}]{stacks-project}).
\begin{equation}\label{eq:pushout}
\begin{tikzcd}
\dvr\ar[d]\ar[r] & \mathcal R_\dvr \ar[d] \\
\mathcal Z_\dvr\ar[r] & \cC_\dvr
\end{tikzcd}
\end{equation}
Recall that the cuspidal curve $\hC_K$ can be characterised as the pushout of the following diagram:
\bcd
2K\ar[r]\ar[d] & \mathcal R_K\ar[d] \\
K\ar[r] & \hC_K.
\ecd
Since the smooth section $\dvr\to\mathcal R_\dvr$ defines a Cartier divisor, it makes sense to take its double and we can thus define $\hC_\dvr$ by means of the similar diagram:
\bcd
2\dvr\ar[r]\ar[d] & \mathcal R_\dvr\ar[d] \\
\dvr\ar[r] & \hC_\dvr
\ecd
The morphism $\phi_\dvr\colon\cC_\dvr\to\hC_\dvr$ extending $\phi_\eta$ is then defined by exploiting the description of $\cC_\dvr$ as a pushout \eqref{eq:pushout}, and the morphisms $\id\colon\mathcal R_\dvr\to \mathcal R_\dvr$ and $\pr_{\mathcal Z_\dvr}\colon\mathcal Z_\dvr\to \dvr$ to the upper right and bottom left corners of the last diagram respectively.
\item If $\cC_\dvr$ does smooth the elliptic tail, then $\phi_\eta$ is an isomorphism. We may assume that $\dvr$ is the spectrum of a complete DVR with algebraically closed residue field \cite[Theorem 7.10]{LMB}. Then we may pick a smooth section for each rational component of $\cC_0$ and extend them to sections of $\cC_\dvr\to \dvr$ by Grothendieck's existence theorem; let us denote by $\Sigma$ the Cartier divisor that is the sum of all such sections. Let $Z$ be the elliptic tail in the central fiber; then we claim that $\omega_{\cC_\dvr/\dvr}(Z)\otimes\mathcal O_{\cC_\dvr}(3\Sigma)$ is $\pi_\dvr$ semi-ample, ample on the generic fiber, and gives the contraction of the elliptic tail to the cusp in the central fiber. We shall not prove the claim here, since this is the core of the second approach.
\end{enumerate}
Finally observe that the map is bijective by construction and $\MM$ is normal, hence $\pi_{|\mathcal X}\colon\mathcal X\to \MM$ is an isomorphism by Zariski's main theorem (as in \cite[\href{http://stacks.math.columbia.edu/tag/082I}{Tag 082I}]{stacks-project}).
 \end{proof}

>>>>

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defprop}[thm]{Definition-Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem*{teo*}{Theorem}
\newtheorem{ipotesi}{ipotesi}
\newtheorem*{nota}{Nota}
\newtheorem{claim}{Claim}
\newtheorem{question}[thm]{Question}

@@

\begin{dfn}
Let $\mathfrak M_{1,n}^{\operatorname{wt}=d,\text{st}}(1)$ be the stack of \emph{at worst cuspidal} projective connected, reduced, $n$-marked, projective curves of arithmetic genus $1$ that are weighted-stable with total weight $d$, i.e. the weight is nonnegative, and every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component of weight $0$ has at least \emph{two} special points).
\end{dfn}
\begin{remark}
 It follows from the miniversal deformation of the cusp that being at worst cuspidal (i.e. having only nodes and cusps as singolarity) is an open condition on the base of any family of curves.
\end{remark}

We shall denote by $\hat{\pi}\colon \widehat{\cC}\to \MM_{1,n}(1)$ the universal curve (and use similar notation for its pullbacks). We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$ of $\widehat{\cC}\to \mathfrak M_{1,n}(1)$ with universal line bundle $\widehat{\mathcal{L}}$ subject to similar stability conditions to the one spelled out for nodal curves. We shall often use shorthand notation $\widehat{\MM}$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}(1)$ and $\widehat{\mathfrak P}$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$. Let $\hat{\PtoM}$ denote the projection $\widehat{\mathfrak P}\to \widehat{\MM}$.

We sometimes work on $\MM_{1,n}^{\rm{div}}$, parametrising families of nodal curves with a relative Cartier divisor, which we may ask to be weighted stable in order to make this space compatible with $\MM$ and $\mathfrak P$ above. More precisely $\MM_{1,n}^{\rm{div}}$ can be thought of as the open inside $C(\pi_*\mathcal L)=\Spec_\pP(\R^1\pi_*\mathcal L)$ (see \cite{CLpfields} and Section \ref{section:p-fields} below), where the section of $\mathcal L$ is not $0$ on any irreducible component of the curve. Alternatively this is the moduli functor of a prestable curve with a line bundle and a section up to scalar, which can be thought of as the hom-stack $\Hom_{\MM_1}(\cC,[\Aaff^1/\Gm])$; then we pick the connected component where the line bundle has total degree $d$, and the open substack obtained by requiring weighted stability and the section not to vanish on any irreducible component of the curve. The morphism $\MM_{1,n}^{\rm{div}}\to\mathfrak P$ is obtained by declaring $\mathcal L_{\MM_{1,n}^{\rm{div}}}:=\OO_{\cC}(\pazocal D)$, where $\pazocal D$ is the universal Cartier divisor. $\MM_{1,n}^{\rm{div}}(1)$ is defined similarly starting with weighted $1$-stable curves.

We shall always denote the spectrum of a discrete valuation ring (DVR) by $\dvr$, with closed point $0$ and generic point $\eta$.

@@

\section{From $p$-fields to the quintic threefold}\label{section:p-fields}
We introduce the moduli space $\Mone{1}{\PP^4}{d}^p$ of $1$-stable curves with $p$-fields, we endow it with a cosection-localised $0$-dimensional virtual class supported on a proper substack (depending on the polynomial $\w$), and we show that its degree coincides with the genus $1$ cuspidal invariants of the quintic threefold $X$ up to a sign. This is a word-by-word repetition of the arguments in \cite{CLpfields} once noticed that they carry over to the situation where we work with a family of Gorenstein (not necessarily nodal) curves. Our aim in this section is to improve the legibility of this paper by providing the non-expert reader with a r\'{e}sum\'{e} of some of the key ideas contained in \cite{CLpfields}; it can otherwise be skipped.
\subsection{Moduli of sections}
\begin{dfn}
 Let $B$ be an algebraic stack and let $\pi\colon \cC\to B$ be a flat proper morphism of finite presentation which is representable by algebraic spaces and whose geometric fibers are reduced l.c.i. curves. Let $\mathcal Z$ be an algebraic stack smooth representable and quasi-projective over $\cC$. The \emph{cone of sections} of $\mathcal Z$ over $\cC$ is a $B$-stack $\mathfrak S$ defined by:
 \[\mathfrak{S}(S\to B)=\left\{\text{sections of}\ \mathcal Z_S\to\cC_S\right\};\]

@@

\begin{lem}\label{lemma:def_X}
 There is a locally closed substack $\X\subseteq\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ representing morphisms $C\to \widehat{C}$ that contract elliptic tails of weight $0$ to cusps and are weight-preserving isomorphisms everywhere else. 
\end{lem}

\begin{lem}\label{lamma:projection_iso}
 The first projection $\pr_1\colon \operatorname{Mor}_{\MM\times\hM}(\cC,\hC)\to\MM$ restricted to $\X$ is an isomorphism with $\MM$.
\end{lem}

\begin{proof}\ref{lemma:def_X}
Recall that $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ is an algebraic stack; in fact the map to $\MM\times\hM$ is representable by algebraic spaces \cite{OlssonHOM}. Let $\phi\colon\cC\to\hC$ denote the universal morphism. We now proceed to construct $\X$ as a locally closed substack in the space of morphisms.

@@

\item[Step 1] We first lift the problem to the level of Picard stacks in order to transmute the property of being weight-preserving into the more manageable one of preserving the line bundles. Recall the notation:
 \[\PtoM\colon\mathfrak P=\mathfrak{Pic}^{tot deg=d,st}_{1,n}\to\MM,\quad \hat\PtoM\colon\hP=\mathfrak{Pic}^{tot deg=d,st}_{1,n}(1)\to\hM\]
  for the Picard stacks of $\pi$ and $\hat \pi$ respectively, with universal line bundles $\L$ and $\hL$. The projections $\PtoM$ and $\hat \PtoM$ are defined by taking the multi-degree of a line bundle. We can now look at the algebraic stack $\operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$ with universal morphism
 $\Phi$ and natural projection $\Pi$ to $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$. There exists a locally closed substack $\pazocal{Y}'\subseteq \operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$ representing those morphisms that \emph{preserve the line bundles}. Indeed, given a smooth chart $S\to\operatorname{Mor}_{\mathfrak{P}\times\hP}(\cC,\hC)$, the locus of $s\in S$ where $\Phi_s^*\hL_s\cong\mathcal{L}_s$ is nothing but the locus $T$ where the two sections $\mathcal{L}_S$ and $\Phi_S^*\hL_S$ of $\mathfrak{P}(S)\to\MM(S)$ are isomorphic. In other words we are looking at the fiber product 
  \[\xymatrix{T\ar[r]\ar[d]&\mathfrak{P}\ar[d]^{\Delta}\\
  S\ar[r]&\mathfrak{P}\times_{\MM}\mathfrak{P}}\]
Being $\mathfrak{P}\to\MM$ representable by locally separated algebraic spaces \cite[Theorem~8.3.1]{neron}, $\Delta$ is a quasi-compact locally closed immersion \cite[\href{http://stacks.math.columbia.edu/tag/04YU}{Tag 04YU}]{stacks-project}, so in particular $T\subseteq S$ is locally closed.

\item[Step 2] 
Furthermore there is a closed substack $\pazocal{Y}\subseteq \pazocal{Y}'$ representing \emph{surjective morphisms that preserve the markings}.

@@

\item[Step 3] We may now get back to $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$. Let $\X'$ be the image of $\pazocal{Y}$ under $\Pi$. This is a constructible substack of $\operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$ by Chevalley's theorem \cite[Theorem 5.9.4]{LMB}. Recall that to show that a constructible set is open (respectively closed) it is enough to check that it contains all the generisations of its points (respectively all the specialisations) \cite[\href{http://stacks.math.columbia.edu/tag/0DQN}{Tag 0DQN}\href{http://stacks.math.columbia.edu/tag/0903}{Tag 0903}]{stacks-project}. Finally, under Noetherian assumptions, two points related by specialisation/generisation are contained in the image of a DVR scheme \cite[\href{http://stacks.math.columbia.edu/tag/054F}{Tag 054F}]{stacks-project}.

@@

\begin{enumerate}
\item\emph{$\phi$ contracts only components of weight $0$.} We show that this is open. Assume that $\dvr$ is a DVR scheme with closed point $0$ and generic point $\eta$, and we are given $\dvr\to\X'$ such that $\phi_0\colon\cC_0\to \hC_0$ does not contract any component of positive weight. Suppose there exists an irreducible component $D_\eta\subseteq \cC_{\eta}$ of positive weight $d_D$ which is contracted by $\phi_{\eta}.$ The contracted locus, i.e. $\left\{c\in \cC_{\dvr} | \dim_c\phi^{-1}(\phi(c))\geq 1\right\}$, is closed by semicontinuity of fiber dimension, hence it contains all the components $D_i\subseteq \cC_0$ to which $D_\eta$ specialises. At least one of them has positive weight, since the sum of their weights is $d_D$, which is a contradiction.

\item\emph{$\phi$ has degree 1 on every non contracted component} or, equivalently, for every $S\to\X'$ there is an $S$-dense open in $\hC_S$ such that the restriction of $\phi_S$ to its preimage is an isomorphism.
This is an open and closed condition; we show it is open. Let $\dvr$ be a DVR scheme as above and assume that $\phi_0$ satisfies the property. Since $\phi_\dvr$ is proper, we may consider 
\[\phi_{\dvr,*}[\cC_{\dvr}]=\sum n_i[\hC_{\dvr,i}]\in A_2(\hC_\dvr).\] Applying Gysin pull-back to $0$ (which is a regular closed point of the base) \cite[Prop. 10.1(a)]{FUL}, we see that all the $n_i$'s are 1 for those $\hC_i$'s such that $0^![\hC_i]\neq 0$. On the other hand there is no irreducible component of $\hC$ supported on $\hC_\eta$.

@@

\item[Step 4] We have imposed that, if $\phi$ contracts a subcurve $E$ of the fiber, it must have weight $0$. Since the fibers of $\hC$ only have nodes and cusps as singularities, and the markings are required to be smooth points, we observe that $E$ must be unmarked and have arithmetic genus $1$ by weighted stability; furthermore $\left|E\cap\overline{C\setminus E}\right|\leq 2$, i.e. $E$ is either an elliptic tail or an elliptic bridge. There are basically two possibilities:
\begin{enumerate}
\item $\phi$ contracts an elliptic tail to a cusp and is an isomorphism everywhere else, or there is no elliptic tail to start with and $\phi$ is an isomorphism;
\item the elliptic tail/bridge is contracted to a smooth point/node, then a non-separating node or a cusp must appear somewhere else in the fiber of $\hC$ in order to preserve the arithmetic genus.
\end{enumerate} 
We want to avoid the second scenario, so we define the open substack $\X\subseteq \X'$ as follows.
Given $\mathcal{C}_S\to\hC_S\in\X'(S)$, let $U\subseteq \hC$ be the \emph{maximal} $S$-dense open subset such that $\phi_S|_{\phi_S^{-1}(U)}\colon\phi_S^{-1}(U)\to U$ is an isomorphism, and $V$ its closed complement in $\hC_S$. 
Then $\X$ is the open locus \cite[\href{http://stacks.math.columbia.edu/tag/055G}{Tag 055G}]{stacks-project} where the fibers of $\pi|_{\phi^{-1}(V)}\colon \phi^{-1}(V)\to S$ are geometrically connected.
\end{description} 
This concludes the construction of the locally closed substack $\X\subseteq \operatorname{Mor}_{\MM\times\hM}(\cC,\hC)$.
 \end{proof}

 \begin{proof}\ref{lamma:projection_iso}
 This result will follow from an application of Zariski's Main Theorem for algebraic spaces.
First we claim that the projection $\pr_{1|\X}\colon\X\to \MM$ is \emph{representable by algebraic spaces}: by \cite[\href{http://stacks.math.columbia.edu/tag/04Y5}{Tag 04Y5}]{stacks-project} we only need to check that it is faithful, and by \cite[Theorem 2.2.5]{CONR} it is enough to look at geometric points. Hence we need to say that, given $\phi\colon C\to \widehat{C}$ a $K$-point of $\X$, we have
$\operatorname{Aut}(\phi)\subseteq\operatorname{Aut}(C)$. Recall that automorphisms of $\phi$ are commutative diagrams:
\bcd C\ar[r,"\phi"]\ar[d,"\psi"] & \widehat{C}\ar[d,"\hat\psi"] \\ C\ar[r,"\phi"] & \widehat{C} \ecd
 Now $\hat\psi$ is determined by $\psi$ due to how $\phi$ behaves (contracting elliptic tails and being an isomorphism everywhere else).

Secondly $\pr_{1|\X}$ is \emph{proper}: this can be seen using the valuative criterion:
\bcd
\eta'=\operatorname{Spec}(K')\ar[r]\ar[d] & \eta=\operatorname{Spec}(K)\ar[r]\ar[d] & \X \ar[d] \\
\dvr'=\operatorname{Spec}(R')\ar[r]\ar[urr,dashed,"\exists?", left] & \dvr=\operatorname{Spec}(R)\ar[r] & \MM
\ecd
Let $\pi \colon \cC_\dvr\to \dvr$ be the family of nodal curves on $\dvr$; there are three cases to consider:
\begin{enumerate}[label=(\alph*)]
\item the central fiber contains no elliptic tail, then the same is true for $\cC_\eta$, hence $\phi_\eta$ is an isomorphism. We can extend $\phi_\eta$ as follows:
\bcd
\cC_\eta\arrow{r}{\sim}[swap]{\phi_\eta}\ar[d,hook,"\iota"] & \hC_\eta\ar[d,hook,"\iota\circ\phi_\eta^{-1}"] \\
\cC_\dvr\ar[r,"\id_\cC"] & \cC_\dvr=:\hC_\dvr
\ecd
Another extension $\phi'\colon \cC_\dvr\cong \hC_\dvr$ would be isomorphic to the previous one via:
\bcd
\cC_\dvr\ar[r,"\phi'"]\ar[d,"\id"] &\hC_\dvr\ar[d,"(\phi')^{-1}"] \\
\cC_\dvr\ar[r,"\id"] & \cC_\dvr
\ecd

\end{enumerate}
If instead $\cC_0$ has got an elliptic tail, then we have two possibilities:
\begin{enumerate}[label=(\alph*)]
\setcounter{enumi}{1}
\item $\cC_{\eta}$ has got an elliptic tail as well; that is the image of $\dvr\to\MM$ is contained in the boundary, so we can find a lift
\bcd
 & \MM_{1,1}\times\MM_{0,1+n}^{\rm{wt}=d,\rm{st}}\ar[d] & \\
 \dvr\ar[r]\ar[ur,dashed] & \mathfrak D_{\{1,\emptyset\},\{0,n\}}\ar[r,hook] &\MM.
\ecd
 Then $\cC_\dvr$ is the pushout of a family of rational curves $\mathcal R_\dvr$ and a family of genus one curves $\mathcal Z_\dvr$:
%The map $S\to\partial \mathfrak M_{1,n}$ lifts to the normalisation of the boundary, i.e. $\mathfrak M_{0,n_0+1}\times\mathfrak M_{1,n_1+1}$, which gives two families of pointed curves on $S$; there is a map from the pushout to $\mathcal C_S$ by the universal property, and it is proper bijective and flat (crit\`{e}re de platitude par fibres \cite[\href{http://stacks.math.columbia.edu/tag/039A}{Tag 039A}]{stacks-project}).
\begin{equation}\label{eq:pushout}
\begin{tikzcd}
\dvr\ar[d]\ar[r] & \mathcal R_\dvr \ar[d] \\
\mathcal Z_\dvr\ar[r] & \cC_\dvr
\end{tikzcd}
\end{equation}
Recall that the cuspidal curve $\hC_K$ can be characterised as the pushout of the following diagram:
\bcd
2K\ar[r]\ar[d] & \mathcal R_K\ar[d] \\
K\ar[r] & \hC_K.
\ecd
Since the smooth section $\dvr\to\mathcal R_\dvr$ defines a Cartier divisor, it makes sense to take its double and we can thus define $\hC_\dvr$ by means of the similar diagram:
\bcd
2\dvr\ar[r]\ar[d] & \mathcal R_\dvr\ar[d] \\
\dvr\ar[r] & \hC_\dvr
\ecd
The morphism $\phi_\dvr\colon\cC_\dvr\to\hC_\dvr$ extending $\phi_\eta$ is then defined by exploiting the description of $\cC_\dvr$ as a pushout \eqref{eq:pushout}, and the morphisms $\id\colon\mathcal R_\dvr\to \mathcal R_\dvr$ and $\pr_{\mathcal Z_\dvr}\colon\mathcal Z_\dvr\to \dvr$ to the upper right and bottom left corners of the last diagram respectively.
\item If $\cC_\dvr$ does smooth the elliptic tail, then $\phi_\eta$ is an isomorphism. We may assume that $\dvr$ is the spectrum of a complete DVR with algebraically closed residue field \cite[Theorem 7.10]{LMB}. Then we may pick a smooth section for each rational component of $\cC_0$ and extend them to sections of $\cC_\dvr\to \dvr$ by Grothendieck's existence theorem; let us denote by $\Sigma$ the Cartier divisor that is the sum of all such sections. Let $Z$ be the elliptic tail in the central fiber; then we claim that $\omega_{\cC_\dvr/\dvr}(Z)\otimes\mathcal O_{\cC_\dvr}(3\Sigma)$ is $\pi_\dvr$ semi-ample, ample on the generic fiber, and gives the contraction of the elliptic tail to the cusp in the central fiber. We shall not prove the claim here, since this is the core of the second approach.
\end{enumerate}
Finally observe that the map is bijective by construction and $\MM$ is normal, hence $\pi_{|\X}\colon\X\to \MM$ is an isomorphism by Zariski's main theorem (as in \cite[\href{http://stacks.math.columbia.edu/tag/082I}{Tag 082I}]{stacks-project}).
 \end{proof}
